---
title: "Robustly subtyping patients"
author: "Paul Agapow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Robustly subtyping patients}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
  %\VignetteEncoding{UTF-8}
---

# devtools::build_vignettes()

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Precision (or personalised or stratified medicine) means giving "the right
treatment for the right patient given at the right time". This in turn
requires that patients be *typed*: assigned to a class that goes beyond
clinical presentation (their appearance) and reflects the particular
mechanisms underlying their disease.

Typing is thus broadly a clustering and classification problem, but
identifying and validating these subtypes can be complex. It is often
unclear which methods work better. In the absence of ground truths, it may be unclear if they work at all. Where there are ground truths, 

diff resolution split / lump
different apis
mulitomics


## Features

This package
presents a number of tools for improved and more consistent subtyping,
including the generation of synthetic data and validation via internal
measures and external ground truths. 

unfication

* Functions for quality control and preprocessing of data
* Metrics and visualisations for assessing the results of subtyping
* Frameworks for exploration and tuning of various parameters
* Several well-understood datasets for testing and validating approaches
* Access to and uniform interfaces for several popular methods


## Conventions

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
