---
title: "A regular worflow of subtyping"
author: Romain Guédon, Paul Agapow
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
number_sections: true
vignette: >
  %\VignetteIndexEntry{Regular workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.retina = 2,
  fig.width = 6,
  cache.lazy = FALSE
)
```

## Aim

Here we'll describe how to use `subtypr` to perform a robust analysis of the 
glioblastoma multiforme (GBM) multi-omic dataset from TCGA that is provided with
the package.

```{r attach_packages, message=FALSE, warning=TRUE}
library(subtypr)
```


## Data exploration

```{r load_gbm}
data_list <- gbm_data$data_list
survival_data <- gbm_data$survival
names(data_list)
ncol(data_list$mrna)
ncol(data_list$methylation)
ncol(data_list$mirna)
```


The GBM dataset contains 273 patients on 3 layers: mRNA, methylation and miRNA with 
12042, 22833 and 534 features respectively.

To have more insight into the data, we'll use `check_distribution()`:

On methylation data:

```{r check_distribution_methylation}
check_distribution(data_list$methylation, thickness = 0.1, split_figures = TRUE)
```

On mRNA data:

```{r check_distribution_mrna}
check_distribution(data_list$mrna, thickness = 0.1, split_figures = TRUE)
```

On miRNA data:

```{r check_distribution_mirna}
check_distribution(data_list$mirna, thickness = 0.3, split_figures = TRUE)
```

## Features selection

There is a lot of features in our methylation data but a majority of them have a variance close to 0. Therefore, we have to 
make a selection to avoid considering features with low variance. To do so, we provide `select_features()` to perform variance-cutoff, MAD-cutoff or PCA.

Here we use the variance-cutoff. 

```{r select_features_var}
new_methylation <- select_features(data = data_list$methylation, method = "variance", cutoff = 0.05)
new_mirna <- select_features(data = data_list$mirna, method = "variance", cutoff = 0.04)
new_mrna <- select_features(data = data_list$mrna, method = "variance", cutoff = 0.5)
```

```{r check_fs_methylation}
check_distribution(new_methylation, thickness = 0.1, split_figures = TRUE)
```

Re-list the data into a new list:

```{r fs_data_list}
fs_data_list <- list(methylation = new_methylation, mirna = new_mirna, mrna = new_mrna)
```

## Apply one subtyping method: Similarity Network Fusion.

We've embedded several methods to perform multi-omics data integration in this package.

To see all the available methods, use:

```{r methods_available}
names(get_method())
```
Here, we'll apply Similarity Network Fusion (SNF). To have more information one the method, use the documentation. For example: `?subtype_snf`. 
Here, we set an arbitraty number of cluster here and we use default parameters of SNF.

Note that to select parameters values, we provide a tuning function (`?tuning`) that tests a grid of parameters and pick up the parameters with the best score (for the selected metric).

```{r}
result_snf <- subtype_snf(data_list = fs_data_list, cluster_number = 2, K = 20, alpha = 0.5, t = 20)
```

All the built-in methods have a similar output structure: the predicted "partition", a character that indicates the data to use for evaluation (e.g. affinity matrix) with standard metrics: "element_for_metric" and other elements:

```{r}
names(result_snf)
```

### Internal metric: silhouette index

Here, we don't have a ground-truth. Therefore, we only have internal metrics to have an idea of the performance of the method. To chose the correct internal metric, see `?metrics_list` to have more information on the built-in metrics.
Here we chose the silhouette index for affinity matrix:

```{r}
silx <- silhouette_affinity(result_snf$partition, result_snf$affinity_fused)
plot(silx)
```


### Survival analysis 

We also have the survival data of these patients. We can see if there is a significant difference 
between the survival curves of the predicted subtypes.

To do so, we use `analyze_survival()`:

```{r analyze_survival, fig.height=4}
analyze_survival(survival_time = survival_data$Survival, 
                 death_status = survival_data$Death, 
                 patients_partition = result_snf$partition)
```

We can see that the subtypes have a significant (5.7e-04) separation between their survival curve.

### Some visualization

We also provide a vizualisation of the results of type affinity matrix sorted by cluster:

```{r plot_affinity, fig.width=5, fig.height=5}
plot_affinity_matrix(result_snf$affinity_fused, result_snf$partition)
```





