---
title: "Tuning parameters"
author: "Romain Guédon, Paul Agapow"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tuning parameters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

How to tune the parameters of a method like Similarity Network Fusion (SNF) ?

SNF requires a number of cluster to apply spectral clustering. It also requires hyperparameters: K, alpha and t. The authors give range of values for alpha but not for K nor t. How to pick the good parameters then ?

We provide `tuning()` to allow a grid search where the performance of the method is assessed by a panel of built-in
metrics.
The idea of `tuning()` is that given:
 * a method, a built-in one or a new one from the user, specified in `method`
 * a set of values for each parameter, specified in `grid_support`
 * a metric, specified in `metric`
 * a partition, specified in `true_partition`, if the selected metric needs one (external metric)

the function will test all the combinations of parameters using parallel computation (if `parallel = TRUE`) and
select the set of parameters with the best value for the selected metric.

Note that `tuning()` can be really computationnaly expensive if the generated combinations by `grid_support` are too large
or if the data itself it too big. Therefore we advise to start exploring hyperparameters with a reasonable step between values (using `seq` for example):

Note also that sometimes there is method-specific tools for tuning the number of cluster. However, these methods also requires the other parameters to be set and because we don't know how to set them, we don't know if the "optimal" number of
cluster suggested can be meaningless. 


```{r library}
library(subtypr)
library(ggplot2)
```


Let's set the grid support:

```{r grid_support}
# The names of the list have to match the names of the method parameters.
grid_support <- list(
  cluster_number = 2:5,
  K = seq(10, 25, 3),
  alpha = seq(0.3, 0.8, 0.1),
  t = c(10, 20)
)
```

We'll use an internal metric to perform the tuning: `metric = "asw_affinity"`. "asw_affinity" is the average silhouette width of the clusters (adapted for affinity matrices instead of distance matrices).

```{r tuning_snf, eval=FALSE}
result_tuning <- tuning(
  data_list = breast_cancer_data$data_list,
  method = "subtype_snf",
  grid_support = grid_support,
  metric = "asw_affinity",
  true_partition = NULL,
  parallel = T,
  ncores = 7,
  plot = F,
  save_results = T,
  file_name = "./tuning_custom_cache/result_tuning_1.RData"
)
```

```{r load, include=FALSE}
load("./tuning_custom_cache/result_tuning_1.RData")
result_tuning <- tuning_result$tuning_result_list
rm(tuning_result)
```

Use `$all_metric_values` in your tuning result to have an overview of the values of the metric for all the parameters:

```{r plot_the_result, fig.width=6, fig.height=4}
plot(result_tuning$all_metric_values, ylab = "Metric values", main = "Metric values of all possibilities generated")
```

Let's compare the performance between default parameters and internally (without groundtruth) tuned parameters:

```{r without_tuning}
no_tuning <- subtype_snf(data_list = breast_cancer_data$data_list, cluster_number = 2)
```


```{r overview_metric, fig.width=8, fig.height=3}
table_overview <- rbind(
  overview_metrics(result_tuning$method_res,
    internal_metrics = "asw_affinity",
    true_partition = breast_cancer_data$partition,
    print = F
  ),
  overview_metrics(no_tuning,
    internal_metrics = "asw_affinity",
    true_partition = breast_cancer_data$partition,
    print = F
  )
)


tuned <- as.factor(c(rep("yes", 8), rep("no", 8)))
ggplot(table_overview) + aes(x = metric, y = value, fill = tuned) + geom_col(position = position_dodge())
```

We can notice a slight improvement in some metrics but SNF is not very sensitive to tuning. 



